---
title: "Untitled"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
```

DE analysis (by DESeq2)
=====================================

```{r}
design.table <- fread(paste0("perl -F':|,' -lanE'BEGIN{say qq{SampleID\tcondition}} $fuck = shift @F; say qq{$_\t$fuck} for @F' ", type.list$Design), header = TRUE)
des.count <- fread(type.list$RSEM)
rna.id <- des.count[,1]
rna.type <- des.count[,2]
des.count <- as.data.frame(des.count[, round(.SD), .SDcols = -(1:2), with = TRUE])
rownames(des.count) <- rna.id[[1]]
coldata <- as.data.frame(design.table[,2])
rownames(coldata) <- names(des.count)
dds <- suppressWarnings(DESeq2::DESeqDataSetFromMatrix(countData = des.count,
                              colData = coldata,
                              design = ~ condition))
keep <- rowSums(DESeq2::counts(dds)) >= 10
dds <- dds[keep,]
dds <- DESeq2::DESeq(dds)
resLFC <- DESeq2::lfcShrink(dds, coef=2)
resOrdered <- res[order(resLFC$pvalue),]
vsd <- DESeq2::vst(dds, blind=FALSE)
ntd <- DESeq2::normTransform(dds)
```

Column {.tabset}
-------------------------------------

### MA-plot

```{r}
DESeq2::plotMA(resLFC, ylim=c(-2,2))
```

### Heatmap

```{r}
select <- order(rowMeans(DESeq2::counts(dds,normalized=TRUE)),
                decreasing=TRUE)[1:20]
# df <- as.data.frame(SummarizedExperiment::colData(dds)[,"condition"])
heatmaply::heatmaply(SummarizedExperiment::assay(ntd)[select,], scale = 'row', xlab = "Sample", margins = c(60,100,40,20), row_text_angle = 45, column_text_angle = 60) %>% layout(margin = list(l = 100, b = 50, r = 0))
```

### Correlation heatmap

```{r}
heatmaply::heatmaply(cor(SummarizedExperiment::assay(ntd)[select,]), margins = c(40, 40),
          k_col = 2, k_row = 2,
          limits = c(-1,1)) %>% layout(margin = list(l = 50, b = 50))
```

### Principle Components Analysis

```{r}
pcaData <- DESeq2::plotPCA(vsd, intgroup="condition", returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
p <- ggplot(pcaData, aes(PC1, PC2, color=condition)) +
  geom_point(size=3) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed()
save_plot('pca.tiff', p, base_height = 8.5, base_width = 11, dpi = 300, compression = 'lzw')
save_plot('pca.pdf', p, base_height = 8.5, base_width = 11, dpi = 300)
ggplotly(p)
```

Column
-------------------------------------

### Description

**Title**: Differential expression analysis of novel and known lncRNA from LncPipe(DESeq2)

DESeq takes all transcripts with similar expression values into consideration. To determine overdispersion parameters in different conditions, it is assumed by DESeq that overdispersion is related to mean reads count. Differentially expressed genes are validated by an exact test between means of distribution parameters in two conditions. In this project, DESeq was last updated as DESeq2 (1.6.2).

**What we have done**: `Design.matrix` file and reads count matrix were fed into [edgeR](version information), an R packages for identification differential expressed gene from RNA-seq or Chip-seq experiment. Before exact test statitistics, raw reads count matrix were filtered according to the user-defined parameter `min.expressed.sample`, which can be explained as follows: raw reads count was first normalized in to cpm (Counts Per Million reads) matrix , gene with cpm >1 in more than `min.expressed.sample` numbers were retained into futher analysis. in current version, we only implemented the standard comparison that only two condition were supportted in each analysis. Experimental without repelicates were not supported . Source code of differential expression analysis can be freely checked by https://github.com/bioinformatist/LncPipeReporter/edit/master/inst/rmd/DE.Rmd. Complementary experimental design can be performed seperately besed `kallisto.count.txt` generated by lncPipe, this file can be imported into another software [IDEA](https://github.com/likelet/IDEA), which focus on comprehensive differential expression analysis from expression matrix.

**What dose plots Mean**: In this section, `MA plot `, `PCA analysis` and `Heatmap plot ` were show at right panel.

`MA plots` A MA-Plot can give a quick overview of the distribution of data. The log2–transformed fold change is plotted on the y-axis and the average count (normalized by size factors) is on the x-axis. The genes with adjusted p-value less than the FDR are colored red, while other genes are colored black.

`PCA analysis` shows Principle Components Analysis result of lncRNA expression matrix, two highest PC were extracted as X and Y axis into scatter plot. Points were colored with different condition involved in this analysis. Theoretically, samples from the same library should be closer than those not. This plot canbe used to control the variance introduced by biological/technical replicates or treatment conditions. In our analysis, PCA analysis were performed using all expressed lncRNA and can be deem as a unsupervised clustering analysis method.

`Heatmap` heatmap plot the correlation matrix, all expressed lncRNA are involved and unsupervised hierarchical clustering method were applied for cluster samples based on correlation matrix, which was displayed in a dendrogram in side.

**Reference** 

Michael I Love, Wolfgang Huber and Simon Anders. Moderated estimation of fold change and dispersion for RNA-seq data with DESeq2. Genome Biology. 2014. 15:550

### Differential expressed lncRNAs table

```{r}
resSig <- BiocGenerics::subset(resOrdered, padj < 0.1)
fwrite(as.data.table(resSig), 'DE.csv', row.names = TRUE)
DT::datatable(head(as.data.table(resSig), n = 80L)) %>% DT::formatRound(c('baseMean', 'log2FoldChange', 'lfcSE', "stat", 'pvalue', 'padj'), digits = 2)
